---
title: "mds_assessment"
author: "AllenY"
date: "5/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results='hide'}
library(ggplot2)
library(tidyr)
library(reshape2)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(tidyr)
library(gplots)
```
```{r}
# read in files
mds <- read.csv('MDS.csv', header = T)
mds_full <- mds
colnames(mds)[2] <- c('days')
head(mds)
```

# long format
```{r}
mds_gather <- gather(mds, key = 'rm', value = 'yield', 3:6, factor_key = T)
mds_gather$date <- as.Date(mds_gather$date)
head(mds_gather)
```

# create bins for plating date
weeks_prior: planting date was at least 7 days earlier than 5/1
peak_week: planting date was between 7 days prior than 5/1 or 7 days after 5/1
1_week_late: planting date was between 7 days and 15 days after 5/1
2_week_late: planting date was between 15 days and 22 days after 5/1
weeks_late: planting date was at least 7 days later than 5/1

```{r}
mds_gather <- mds_gather%>%
  mutate(day_range = case_when(days < -7 ~ 'weeks_prior',
                               days >= -7 & days < 7 ~ 'peak_week',
                               days >= 7 & days < 15 ~'1_week_late',
                               days >=15 & days < 22 ~ '2_weeks_late',
                               days >=22 ~ 'weeks_late'))
mds_gather$day_range <- as.factor(mds_gather$day_range)
mds_gather$day_range <- factor(mds_gather$day_range, 
                               levels = c('weeks_prior','peak_week','1_week_late','2_weeks_late', 'weeks_late'))
head(mds_gather)
```

# anova for the rm and day_range
```{r}
aov.rm_dr <- aov(yield ~ rm*day_range, data = mds_gather)
summary(aov.rm_dr)
```
Since day_range is significant, we can single out each day_range and find the best rm within.

# weeks_prior
```{r}
wp <- mds_gather[mds_gather[,'day_range']=='weeks_prior', ]

aov_wp <- aov(yield ~ rm, data = wp)
summary(aov_wp)
```
Mature group's effect was not significant in this case, meaning there is no significant differences
in yield loss between different mature group.

# peak_week
```{r}
pw <- mds_gather[mds_gather[,'day_range']=='peak_week', ]

aov_pw <- aov(yield ~ rm, data = pw)
summary(aov_pw)
```
Mature group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

# 1_week_late
```{r}
wl1 <- mds_gather[mds_gather[,'day_range']=='1_week_late', ]

aov_wl1 <- aov(yield ~ rm, data = wl1)
summary(aov_wl1)
```
Mature group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

# weeks_late
```{r}
wl2 <- mds_gather[mds_gather[,'day_range']=='2_weeks_late', ]

aov_wl2 <- aov(yield ~ rm, data = wl2)
summary(aov_wl2)
```
Mature group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

# weeks_late
```{r}
wls <- mds_gather[mds_gather[,'day_range']=='weeks_late', ]

aov_wls <- aov(yield ~ rm, data = wls)
summary(aov_wls)
```
Mature group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.


```{r}
ggplot(data = mds_gather, aes(x=days, y=yield, color = rm))+
  geom_smooth(method='lm')+
  geom_line()+
  labs(title='Yield through planting date')
```
The straight lines are the regression line for each mature group. According to this dataset, the loss of yield
would increase through time.


```{r}
hm <- as.matrix(mds[,c(3:6)])
rownames(hm) <- mds$days
hm_dev <- hm - colMeans(hm)
heatmap.2(hm_dev, scale = 'none', col = bluered(100), trace = 'none',
          cexRow = 0.7, cexCol = 1, lhei = c(0.5,1.5), lwid = c(0.5,1.5), 
          margins = c(6,5), keysize = 2, xlab = 'Maturity Group',
          ylab = 'Days after 5/1', main = 'Yield deviation according to plating date')

```
Each column represents the yield for each plating date for each maturity group.
Increase in red means the yield are higher than the average. Increase in blue means
the yield is lower than average.
This graph tells us what dates to plant would have higher yield and what dates to plant
would have lower yield.


# best day_range to plant
Since we know day_range is sigificant we should find out which day_range would be the best
and worst in yield.
```{r}
aov_dr <- aov(yield ~ day_range, data = mds_gather)
tk_dr <- TukeyHSD(aov_dr)
tk_dr
```
According to Tukey's post-test:
peak_week is significantly higher than 2_weeks_late and weeks_late in yield. (with a average yield difference of 15.28 B/A)
1_week_late is significantly higher than 2_weeks_late and weeks_late in yield.  (with a average yield difference of 15.85 B/A)
weeks_prior, peak_week, and 1_week_late does not have difference in yield.
weeks_prior, 2_week_late, and weeks_late does not have differnce in yield.

# boxplot
```{r}
ggplot(data = mds_gather, aes(x=day_range, y=yield, color=rm))+
  geom_boxplot()
```


