---
title: "mds_assessment"
author: "AllenY"
date: "5/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results='hide'}
library(ggplot2)
library(tidyr)
library(reshape2)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(tidyr)

```
```{r}
# read in files
mds <- read.csv('MDS.csv', header = T)
mds_full <- mds
colnames(mds)[2] <- c('days')
head(mds)
```

# long format
```{r}
mds_gather <- gather(mds, key = 'rm', value = 'yield', 3:6, factor_key = T)
mds_gather$date <- as.Date(mds_gather$date)
head(mds_gather)
```

# create bins for plating date
weeks_prior: planting date was at least 7 days earlier than 5/1
peak_week: planting date was between 7 days prior than 5/1 or 7 days after 5/1
1_week_late: planting date was between 7 days and 15 days after 5/1
2_week_late: planting date was between 15 days and 22 days after 5/1
weeks_late: planting date was at least 7 days later than 5/1

```{r}
mds_gather <- mds_gather%>%
  mutate(day_range = case_when(days < -7 ~ 'weeks_prior',
                               days >= -7 & days < 7 ~ 'peak_week',
                               days >= 7 & days < 15 ~'1_week_late',
                               days >=15 & days < 22 ~ '2_weeks_late',
                               days >=22 ~ 'weeks_late'))
mds_gather$day_range <- as.factor(mds_gather$day_range)
mds_gather$day_range <- factor(mds_gather$day_range, 
                               levels = c('weeks_prior','peak_week','1_week_late','2_weeks_late', 'weeks_late'))
head(mds_gather)
```

# anova for the rm and day_range
```{r}
aov.rm_dr <- aov(yield ~ rm*day_range, data = mds_gather)
summary(aov.rm_dr)
```
Since day_range is significant, we can single out each day_range and find the best rm within.

# weeks_prior
```{r}
wp <- mds_gather[mds_gather[,'day_range']=='weeks_prior', ]

aov_wp <- aov(yield ~ rm, data = wp)
summary(aov_wp)
```
Mature group's effect was not significant in this case, meaning there is no significant differences
in yield loss between different mature group.

# peak_week
```{r}
pw <- mds_gather[mds_gather[,'day_range']=='peak_week', ]

aov_pw <- aov(yield ~ rm, data = pw)
summary(aov_pw)
```
Mature group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

# 1_week_late
```{r}
wl1 <- mds_gather[mds_gather[,'day_range']=='1_week_late', ]

aov_wl1 <- aov(yield ~ rm, data = wl1)
summary(aov_wl1)
```
Mature group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

# weeks_late
```{r}
wl2 <- mds_gather[mds_gather[,'day_range']=='2_weeks_late', ]

aov_wl2 <- aov(yield ~ rm, data = wl2)
summary(aov_wl2)
```
Mature group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

# weeks_late
```{r}
wls <- mds_gather[mds_gather[,'day_range']=='weeks_late', ]

aov_wls <- aov(yield ~ rm, data = wls)
summary(aov_wls)
```
Mature group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

# boxplot
```{r}
ggplot(data = mds_gather, aes(x=day_range, y=yield, color=rm))+
  geom_boxplot()
```


