---
title: "MDS_dataset"
author: "AllenY"
date: "5/5/2020"
output:
  html_document: default
  pdf_document: default
---
## Question to answer:
Please show the loss of the yields depending on the planting date in the 4 maturity groups (Early4, Medium 4, Late 4, and Early 5).  
Also if you notice anything else about the data that you feel should be analyzed and reported, please do so.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyr)
library(reshape2)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(tidyr)
library(gplots)
```
```{r}
# read in files
mds <- read.csv('MDS.csv', header = T)
mds_full <- mds
colnames(mds)[2] <- c('days')
head(mds)
```

### Create heatmap in R (Same as the graph in excel)
```{r}
hm <- as.matrix(mds[,c(3:6)])
rownames(hm) <- mds$days
hm_dev <- hm - colMeans(hm)
heatmap.2(hm_dev, scale = 'none', col = bluered(100), trace = 'none',
          cexRow = 0.7, cexCol = 1, lhei = c(0.5,1.5), lwid = c(0.5,1.5), 
          margins = c(6,5), keysize = 2, xlab = 'Maturity Group',
          ylab = 'Days after 5/1')

```

Each column represents the yield for each plating date for each maturity group.  
Increase in red means the yield is higher than the Maturity Group's average. Increase in blue means
the yield is lower than the Maturity Group's average.  
This graph tells us what dates to plant would have higher yield and what dates to plant
would have lower yield.  


```{r}
# long format
mds_gather <- gather(mds, key = 'rm', value = 'yield', 3:6, factor_key = T)
mds_gather$date <- as.Date(mds_gather$date)
head(mds_gather)
```

### Create new variable: day_range
Since the information I have on the time variable wasn't sufficient enough to 
perform proper statistical analysis, I created a new time variable "day_range"- bins
to group similar dates.  

weeks_prior: planting date was at least 7 days earlier than 5/1  
peak_week: planting date was between 7 days prior than 5/1 or 7 days after 5/1  
1_week_late: planting date was between 7 days and 15 days after 5/1  
2_week_late: planting date was between 15 days and 22 days after 5/1  
weeks_late: planting date was at least 22 days later than 5/1  

```{r}
# create bins for planting date
mds_gather <- mds_gather %>%
  mutate(day_range = case_when(days < -7 ~ 'weeks_prior',
                               days >= -7 & days < 7 ~ 'peak_week',
                               days >= 7 & days < 15 ~'1_week_late',
                               days >=15 & days < 22 ~ '2_weeks_late',
                               days >=22 ~ 'weeks_late'))
mds_gather$day_range <- as.factor(mds_gather$day_range)
mds_gather$day_range <- factor(mds_gather$day_range, 
                               levels = c('weeks_prior','peak_week','1_week_late','2_weeks_late', 'weeks_late'))
head(mds_gather)
```

### ANOVA for variable rm and day_range
```{r}
# anova rm and day_range
aov.rm_dr <- aov(yield ~ rm*day_range, data = mds_gather)
summary(aov.rm_dr)
```
Since day_range is significant, we can single out each day_range and find the best rm within.

### ANOVA for Maturity Group in day_range: weeks_prior
```{r}
# weeks_prior
wp <- mds_gather[mds_gather[,'day_range']=='weeks_prior', ]

aov_wp <- aov(yield ~ rm, data = wp)
summary(aov_wp)
```
Maturity group's effect was not significant in this case, meaning there is no significant differences
in yield loss between different mature group.

### ANOVA for Maturity Group in day_range: peak_week
```{r}
# peak_week
pw <- mds_gather[mds_gather[,'day_range']=='peak_week', ]

aov_pw <- aov(yield ~ rm, data = pw)
summary(aov_pw)
```
Maturity group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

### ANOVA for Maturity Group in day_range: 1_week_late
```{r}
wl1 <- mds_gather[mds_gather[,'day_range']=='1_week_late', ]

aov_wl1 <- aov(yield ~ rm, data = wl1)
summary(aov_wl1)
```
Maturity group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.


### ANOVA for Maturity Group in day_range: 2_weeks_late
```{r}
# 2_weeks_late
wl2 <- mds_gather[mds_gather[,'day_range']=='2_weeks_late', ]

aov_wl2 <- aov(yield ~ rm, data = wl2)
summary(aov_wl2)
```
Maturity group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.


### ANOVA for Maturity Group in day_range: weeks_late
```{r}
# weeks_late
wls <- mds_gather[mds_gather[,'day_range']=='weeks_late', ]

aov_wls <- aov(yield ~ rm, data = wls)
summary(aov_wls)
```
Maturity group's effect was not significant in this case as well, meaning there is no significant differences
in yield loss between different mature group.

### Plot each Maturity group's daily yield against planting day
```{r, warning=FALSE}
ggplot(data = mds_gather, aes(x=days, y=yield, color = rm))+
  geom_smooth(method='lm')+
  geom_line()+
  labs(title='Yield through planting date')
```
  
The straight lines are the regression lines for each mature group.  
According to this dataset, the loss of yield would increase through time.


### Best day_range to plant
Since we know day_range effect is sigificant, we can find out which day_range would be the best
and worst in yield.  
```{r}
aov_dr <- aov(yield ~ day_range, data = mds_gather)
tk_dr <- TukeyHSD(aov_dr)
tk_dr
```
According to Tukey's post-hoc test:  
peak_week is significantly higher than 2_weeks_late and weeks_late in yield. (With an average yield difference of 15.28 B/A)  
1_week_late is significantly higher than 2_weeks_late and weeks_late in yield.  (With an average yield difference of 15.85 B/A)  
weeks_prior, peak_week, and 1_week_late does not have difference in yield.  
weeks_prior, 2_week_late, and weeks_late does not have differnece in yield.  

### Boxplot for each day_range
```{r}
# boxplot
ggplot(data = mds_gather, aes(x=day_range, y=yield, color=rm))+
  geom_boxplot()+
  labs(title = 'Differences in Yield across day_range')+
  ylab('Yield')
```

## Summary
According to this dataset, I conclude that Maturity Group does NOT have significant effect in all cases- meaning no 
matter what Maturity Group the plants were, yields do not have any difference.  
However, the best planting day would be anytime prior to 15 days after 5/1- it would have sigificantly 
higher yields overall compared to other planting day.  
In addtion, planting anytime later than 15 days after 5/1 would have sigificantly higher yield loss.  
A reasonable explaination would be because soybeans are highly-sensitive to the duration of sunlight,
the longer the time plants were able to recevie light, the better the yield would be. I would assume that
this dataset contains multuple locations, and thus maturity group's effect was diluted.

  
  
  



