---
title: "population_dataset"
author: "AllenY"
date: "5/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  lines <- options$output.lines
  if (is.null(lines)) {
    return(hook_output(x, options))  # pass to default hook
  }
  x <- unlist(strsplit(x, "\n"))
  more <- "..."
  if (length(lines)==1) {        # first n lines
    if (length(x) > lines) {
      # truncate the output, but add ....
      x <- c(head(x, lines), more)
    }
  } else {
    x <- c(more, x[lines], more)
  }
  # paste these lines together
  x <- paste(c(x, ""), collapse = "\n")
  hook_output(x, options)
})
```

# Questions to answer:
### Which is the best population per variety
### Which is the best population per RM ( maturity group)
### Which is the best population per RM in different environment ( High environment > 75 B/A – Medium Environment  60 – 75 B/A – low environment < 60 B/A)

```{r, message=FALSE, warning=FALSE}
# load necessary packages
library(ggplot2)
library(tidyr)
library(reshape2)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(tidyr)
```

```{r}
# read in files
pop <- read.csv('Population.csv')
pop <- pop[1:9]
pop$Population <- as.factor(pop$Population)
head(pop)
```
## 1. Which is the best population per variety?
In order to answer this question, we need to confirm that there is significant differences between different population-
to say that that particular population(s) is the best for that variety.

### Check for signicant variables
```{r}
# anova population*variety
var_pop <- aov(Yield ~ Population*Variety, data = pop)
summary(var_pop)
```
  
There seems to be no significant effect on yield for population. The interaction effect was not significant as well.
However, Variety showed significant effect on yield.  
Therefore, I decided to find the best population within each variety.

### Best population for DM 3810
```{r}
# ANOVA for population within DM 4210
dm_3810 <- pop[pop[,'Variety']=='DM 3810', ]

aov.pop_dm3810 <- aov(Yield ~ Population, data = dm_3810)
summary(aov.pop_dm3810)
TukeyHSD(aov.pop_dm3810)

ggplot(data = dm_3810, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in DM 3810's Yield between different Population")
```
  
Statiscally speaking, there is no difference between Populations within the same variety. Meaning
no matter which Population DM 3810 were grew, there is no difference in Yield.


### Best population within DM 4210
```{r}
# ANOVA for population within DM 4210
dm_4210 <- pop[pop[,'Variety']=='DM 4210', ]

aov.pop_dm4210 <- aov(Yield ~ Population, data = dm_4210)
summary(aov.pop_dm4210)
TukeyHSD(aov.pop_dm4210)

ggplot(data = dm_4210, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in DM 4210's Yield between different Population")
```
  
Similar result as in DM 3819, statiscally speaking, there is no difference between Populations 
within the same variety. Meaning no matter which Population DM 4210 were grew, there is no difference in Yield.  
However, Population 25 would potentially have higher yield than Population 35.

### Best population within DM 4612
```{r}
# ANOVA for population within DM 4612
dm_4612 <- pop[pop[,'Variety']=='DM 4612', ]

aov.pop_dm4612 <- aov(Yield ~ Population, data = dm_4612)
summary(aov.pop_dm4612)
TukeyHSD(aov.pop_dm4612)

ggplot(data = dm_4612, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in DM 4612's Yield between different Population")
```
  
Similar result, statiscally speaking, there is no difference between Populations 
within the same variety. Meaning no matter which Population DM 4612 were grew, it will have consistent yield.

### Best population within DM 4670
```{r}
# ANOVA for population within DM 4670
dm_4670 <- pop[pop[,'Variety']=='DM 4670', ]

aov.pop_dm4670 <- aov(Yield ~ Population, data = dm_4670)
summary(aov.pop_dm4670)
TukeyHSD(aov.pop_dm4670)

ggplot(data = dm_4670, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in DM 4670's Yield between different Population")
```
  
Similar result, statiscally speaking, there is no difference between Populations 
within the same variety. Meaning no matter which Population DM 4670 were grew, it will have consistent yield.

### Best population within DM 4712
```{r}
# ANOVA for population within DM 4712
dm_4712 <- pop[pop[,'Variety']=='DM 4712', ]

aov.pop_dm4712 <- aov(Yield ~ Population, data = dm_4712)
summary(aov.pop_dm4712)
TukeyHSD(aov.pop_dm4712)

ggplot(data = dm_4712, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in DM 4712's Yield between different Population")
```
  
Similar result, statiscally speaking, there is no difference between Populations 
within the same variety. Meaning no matter which Population DM 4712 were grew, it will have consistent yield.

### Best population within DM 4970
```{r}
# ANOVA for population within DM 4970
dm_4970 <- pop[pop[,'Variety']=='DM 4970', ]

aov.pop_dm4970 <- aov(Yield ~ Population, data = dm_4970)
summary(aov.pop_dm4970)
TukeyHSD(aov.pop_dm4970)

ggplot(data = dm_4970, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in DM 4970's Yield between different Population")
```
  
Similar result, statiscally speaking, there is no difference between Populations 
within the same variety. Meaning no matter which Population DM 4970 were grew, it will have consistent yield.

### Best for population within DM 5.1
```{r}
# ANOVA for population within DM 5.1
dm_5.1 <- pop[pop[,'Variety']=='DM 5.1', ]

aov.pop_dm5.1 <- aov(Yield ~ Population, data = dm_5.1)
summary(aov.pop_dm5.1)
TukeyHSD(aov.pop_dm5.1)

ggplot(data = dm_5.1, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in DM 5.1's Yield between different Population")
```
  
Though the boxplot shows that there seems to be a difference in mean, the sample size
was too small to make a statistical conclustion. Therefore, there's no statistic 
difference in yield between Population.

### Conclusion for the first question
There's no significant difference in yield between Population for each Vareity- meaning
all varieties have consistent yield across Population. A honorable mention would be DM 4210. It
would potentially have higher yield in Population 25 than Population 35.
  
  

## 2. Which is the best population per RM (maturity group)?
Similar approach as to answer the first question. But this time, instead of 
lookning into each variety, each maturity group will be examined.

### Significant test for variables
```{r}
rm_pop <- aov(Yield ~ RM * Population, data = pop)
summary(rm_pop)
```
  
Maturity group have significant difference in yield, while population does not.


### Best Population for Early 4
```{r}
e4 <- pop[pop[,'RM']=='Early 4', ]

aov.pop_e4 <- aov(Yield ~ Population, data = e4)
summary(aov.pop_e4)
TukeyHSD(aov.pop_e4)

ggplot(data = e4, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in Early 4's Yield between different Population")
```
  
There is no significant difference in yield between populations when planting Early 4.

### Best Population for Early 5
```{r}
e5 <- pop[pop[,'RM']=='Early 5', ]

aov.pop_e5 <- aov(Yield ~ Population, data = e5)
summary(aov.pop_e5)
TukeyHSD(aov.pop_e5)

ggplot(data = e5, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in Early 5's Yield between different Population")
```
  
At glance, Population 25 seems to have higher yield than Population 15 and 35 when planting Early 5. 
However, statistically, I am 95% confidence that Population 25 does NOT differ from Population 15 and 35.  
Mainly because the sample size was too small to conlcude that the two population are significantly different.


### Best Population for Late 3
```{r}
l3 <- pop[pop[,'RM']=='Late 3', ]

aov.pop_l3 <- aov(Yield ~ Population, data = l3)
summary(aov.pop_l3)
TukeyHSD(aov.pop_l3)

ggplot(data = l3, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in Late 3's Yield between different Population")

```
  
When planting Late 3, there is no significant difference in Yield between different Population.
  
### Best Population for Late 4
```{r}
l4 <- pop[pop[,'RM']=='Late 4', ]

aov.pop_l4 <- aov(Yield ~ Population, data = l4)
summary(aov.pop_l4)
TukeyHSD(aov.pop_l4)

ggplot(data = l4, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in Late 4's Yield between different Population")
```
  
There is also no significant difference in Yield between different Population when plating Late 4.

### Best Population for Medium 4
```{r}
m4 <- pop[pop[,'RM']=='Medium 4', ]

aov.pop_m4 <- aov(Yield ~ Population, data = m4)
summary(aov.pop_m4)
TukeyHSD(aov.pop_m4)

ggplot(data = m4, aes(x=Population, y=Yield))+
  geom_boxplot()+
  labs(title = "Differences in Medium 4's Yield between different Population")
```
  
There's no significant difference in yield between population when planting Medium 4.

### Conclusion for the second question
There's no significant difference in yield between population for each maturity group- meaning
every maturity group perform the same across all population.
  
  
## 3. Which is the best population per RM in different environment (High environment > 75 B/A – Medium Environment  60 – 75 B/A – low environment < 60 B/A)
For this question, though the approach would be similar as previous, we need to create a new variable to identify
different environment.
```{r}
# create environment
pop <- pop %>%
  mutate(environment = case_when(Yield > 75 ~ 'high',
                                 (Yield <= 75 & Yield >= 60) ~'medium',
                                 Yield < 60 ~ 'low'))
```

### ANOVA test for significant variable
```{r}
env_rm_pop <- aov(Yield ~ environment*RM*Population, data = pop)
summary(env_rm_pop)
```
  
Environment and Population now have significant effect in Yield. 
An interesting detail is that Environment have significant interaction effect on RM and Population-
meaning when the Environment changed, the performance of RM and Population would change
as well.  
We can now proceed to find the best population for each RM in different Environment.

### High environment
```{r}
high <- pop[pop[,'environment']=='high', ]
aov.high_pop_rm <- aov(Yield ~ Population * RM, data = high)
summary(aov.high_pop_rm)
```
  
In high environment, Maturity groups have significantly different yield. However, 
Population does NOT seem to have significant difference in yield. Therefore, to speed up 
the analysis process, I directly compared all means for each combination

```{r, output.lines = 25}
TukeyHSD(aov.high_pop_rm)
```
  
There is no signifcant difference in yield for the populations in each maturity group in high environment.
(See full output in rmd file)

```{r}
# Boxplot for Yields betwen Population for each RM
ggplot(data = high, aes(x=RM, y=Yield, color=Population))+
  geom_boxplot()+
  labs(title = 'Yields betwen Population for each RM in High environment')
```

### Medium environment
```{r}
med <- pop[pop[,'environment']=='medium', ]
aov.med_pop_rm <- aov(Yield ~ Population * RM, data = med)
summary(aov.med_pop_rm)
```
  
In medium environment, none of the variable efffects were significant- meaning no matter what 
population the plants were planted, and no matter what maturity group were used, there is no
significant difference in yield.

### Check to see if there is significant difference within each maturity group
```{r, output.lines = 25}
TukeyHSD(aov.med_pop_rm)
```
  
There is no signifcant difference in yield for the populations in each maturity group in medium environment.
(See full output in rmd file)

```{r}
# Boxplot for Yields betwen Population for each RM
ggplot(data = med, aes(x=RM, y=Yield, color=Population))+
  geom_boxplot()+
  labs(title = 'Yields betwen Population for each RM in Medium environment')
```
  
Again, the graph might seems to tell that there's different performance for different Population, but  
the sample size was not sufficient enough to say the yield are significantly different between each population.

### Low environment
```{r}
low <- pop[pop[,'environment']=='low', ]
aov.low_pop_rm <- aov(Yield ~ Population * RM, data = low)
summary(aov.low_pop_rm)
```
  
In low environment, there is significant difference in between populations

```{r, output.lines = 25}
TukeyHSD(aov.low_pop_rm)
```
  
In low environment, when planting Medium 4, Population 45 and 35 is significantly higher than Population 15.
(See full output in rmd file)

```{r}
# Boxplot for Yields betwen Population for each RM
ggplot(data = low, aes(x=RM, y=Yield, color=Population))+
  geom_boxplot()+
  labs(title = 'Yields betwen Population for each RM in Low environment')
```
  
Though the graph does not seems to have clear difference for Population 15 vs. 35 and Population 15 vs 45, 
a good explaination would be when planting Medium 4 in low environment with Population 15, the yield would have
drastic difference. Therefore, it would be better to select either population 25, 35, or 45 when planting
Medium 4 in low environments.

## Conclusion for question 3
For high and medium environment, there is no population that stands out significantly within each
maturity group- every maturity group perform stably across all population. However, for low environment 
it would be more ideal to select population 25, 35 or 45 when planting Medium 4, since it will not perform 
stable in population 15.









